---
title: "csulb_calibration_test_nov_14"
author: "John Parsons"
date: "11/17/2021"
output: html_document
---

```{r}
knitr::opts_knit$set("C:/Users/Dell/Desktop/MA_analysis")
library(gt)
```

# Comparing between drones
```{r}
library(tidyverse)

comp_dat <- read.csv("data_raw/csulb_tests_nov_14/drone_comp.txt")
View(comp_dat)

comp_dat %>% 
  group_by(test_alt, drone) %>% 
  summarize(pvc = list(mean_se(size_pvc_asl)), board = list(mean_se((size_board_asl)))) %>% 
  unnest(pvc) %>% 
  mutate(se_pvc = y - ymin) %>% 
  rename(pvc = y) %>% 
  select(-c(4,5)) %>% 
  unnest(board) %>% 
  mutate(se_board = y - ymin) %>% 
  rename(board = y) %>% 
  select(-c(5,6))
```

# Submerged test (1.5 m depth)
```{r}
BOI_1_sub_dat <- read.csv("data_raw/csulb_tests_nov_14/BOI_1_submerged_JP_calculations.csv")
View(BOI_1_sub_dat)

BOI_1_sub_dat$test_alt <- as.factor(as.character(BOI_1_sub_dat$test_alt))

BOI_1_sub_dat %>%
  ggplot(aes(x = test_alt, y = size_pvc, fill = method)) +
    geom_boxplot() +
    geom_hline(yintercept = 1.97, linetype = 2, color = "red")
ggsave("plots/calibration_tests/submerged.png", width = 10)

BOI_1_sub_dat %>%
  ggplot(aes(x = test_alt, y = size_pvc_asl, fill = method)) +
    geom_boxplot() +
    geom_hline(yintercept = 1.97, linetype = 2, color = "red")
ggsave("plots/calibration_tests/submerged_asl.png", width = 10)

BOI_1_sub_dat %>%
  ggplot(aes(x = test_alt, y = size_asl_plus_depth, fill = method)) +
    geom_boxplot() +
    geom_hline(yintercept = 1.97, linetype = 2, color = "red")
ggsave("plots/calibration_tests/submerged_asl_plus_depth.png", width = 10)

BOI_1_sub_dat %>%
  pivot_longer(cols = c(5:7), names_to = "height_type", values_to = "length") %>%
  filter(method == "video") %>% 
  ggplot(aes(x = test_alt, y = length, fill = height_type)) +
    geom_boxplot() +
    geom_hline(yintercept = 1.97, linetype = 2, color = "red")
ggsave("plots/calibration_tests/height_comp_submerged.png", width = 10)
```

# Surface test, BOI 1 drone
```{r}
BOI_1_surface_dat <- read.csv("data_raw/csulb_tests_nov_14/BOI_1_surface_JP_calculations.csv")
View(BOI_1_surface_dat)

BOI_1_surface_dat$test_alt <- as.factor(as.character(BOI_1_surface_dat$test_alt))

BOI_1_surface_dat %>%
  ggplot(aes(x = test_alt, y = size_pvc, fill = method)) +
    geom_boxplot() +
    geom_hline(yintercept = 1.98, linetype = 2, color = "red")
ggsave("plots/calibration_tests/BOI_1_surface_pvc.png", width = 10)

BOI_1_surface_dat %>%
  ggplot(aes(x = test_alt, y = size_board, fill = method)) +
    geom_boxplot() +
    geom_hline(yintercept = 2.93, linetype = 2, color = "red")
ggsave("plots/calibration_tests/BOI_1_surface_board.png", width = 10)

BOI_1_surface_dat %>%
  ggplot(aes(x = test_alt, y = size_pvc_asl, fill = method)) +
    geom_boxplot() +
    geom_hline(yintercept = 1.98, linetype = 2, color = "red")
ggsave("plots/calibration_tests/BOI_1_surface_pvc_asl.png", width = 10)

BOI_1_surface_dat %>%
  ggplot(aes(x = test_alt, y = size_board_asl, fill = method)) +
    geom_boxplot() +
    geom_hline(yintercept = 2.93, linetype = 2, color = "red")
ggsave("plots/calibration_tests/BOI_1_surface_board_asl.png", width = 10)

BOI_1_surface_dat %>%
  pivot_longer(cols = c(5:8), names_to = "height_type", values_to = "length") %>%
  filter(method == "video") %>% 
  ggplot(aes(x = test_alt, y = length, fill = height_type)) +
    geom_boxplot() +
    geom_hline(yintercept = c(1.97, 2.93), linetype = 2, color = "red")
ggsave("plots/calibration_tests/height_comps_surface.png", width = 10)
```
# Surface test, BOI 2 drone
```{r}
BOI_2_surface_dat <- read.csv("data_raw/csulb_tests_nov_14/BOI_2_surface_JP_calculations.csv")
View(BOI_2_surface_dat)

BOI_2_surface_dat$test_alt <- as.factor(as.character(BOI_2_surface_dat$test_alt))

BOI_2_surface_dat %>%
  ggplot(aes(x = test_alt, y = size_pvc, fill = method)) +
    geom_boxplot() +
    geom_hline(yintercept = 1.98, linetype = 2, color = "red")
ggsave("plots/calibration_tests/BOI_2_surface_pvc.png", width = 10)

BOI_2_surface_dat %>%
  ggplot(aes(x = test_alt, y = size_board, fill = method)) +
    geom_boxplot() +
    geom_hline(yintercept = 2.93, linetype = 2, color = "red")
ggsave("plots/calibration_tests/BOI_2_surface_board.png", width = 10)

BOI_2_surface_dat %>%
  ggplot(aes(x = test_alt, y = size_pvc_asl, fill = method)) +
    geom_boxplot() +
    geom_hline(yintercept = 1.98, linetype = 2, color = "red")
ggsave("plots/calibration_tests/BOI_2_surface_pvc_asl.png", width = 10)

BOI_2_surface_dat %>%
  ggplot(aes(x = test_alt, y = size_board_asl, fill = method)) +
    geom_boxplot() +
    geom_hline(yintercept = 2.93, linetype = 2, color = "red")
ggsave("plots/calibration_tests/BOI_2_surface_board_asl.png", width = 10)
```

# Comparing BOI 1 to BOI 2
```{r}
BOI_1_vs_2_dat <- bind_rows(BOI_1_surface_dat, BOI_2_surface_dat, .id = "drone")
View(BOI_1_vs_2_dat)

BOI_1_vs_2_dat %>% 
  filter(method == "video") %>%
  ggplot(aes(x = test_alt, y = size_pvc_asl, fill = drone)) +
    geom_boxplot() +
    geom_hline(yintercept = 1.98, linetype = 2, color = "red")
ggsave("plots/calibration_tests/BOI_1_vs_2_pvc_asl.png", width = 10)

BOI_1_vs_2_dat %>% 
  filter(method == "video") %>%
  ggplot(aes(x = test_alt, y = size_board_asl, fill = drone)) +
    geom_boxplot() +
    geom_hline(yintercept = 2.93, linetype = 2, color = "red")
ggsave("plots/calibration_tests/BOI_1_vs_2_board_asl.png", width = 10)

BOI_1_vs_2_dat %>% 
  filter(method == "photo") %>%
  ggplot(aes(x = test_alt, y = size_board_asl, fill = drone)) +
    geom_boxplot() +
    geom_hline(yintercept = 2.93, linetype = 2, color = "red")
```

# Surface test, Love Lab drone
```{r}
LL_surface_dat <- read.csv("data_raw/csulb_tests_nov_14/LL_surface_JP_calculations.csv")
View(LL_surface_dat)

LL_surface_dat$test_alt <- as.factor(as.character(LL_surface_dat$test_alt))

LL_surface_dat %>%
  ggplot(aes(x = test_alt, y = size_pvc, fill = method)) +
    geom_boxplot() +
    geom_hline(yintercept = 1.98, linetype = 2, color = "red")
ggsave("plots/calibration_tests/LL_surface_pvc.png", width = 10)

LL_surface_dat %>%
  ggplot(aes(x = test_alt, y = size_board, fill = method)) +
    geom_boxplot() +
    geom_hline(yintercept = 2.93, linetype = 2, color = "red")
ggsave("plots/calibration_tests/LL_surface_board.png", width = 10)

LL_surface_dat %>%
  ggplot(aes(x = test_alt, y = size_pvc_asl, fill = method)) +
    geom_boxplot() +
    geom_hline(yintercept = 1.98, linetype = 2, color = "red")
ggsave("plots/calibration_tests/LL_surface_pvc_asl.png", width = 10)

LL_surface_dat %>%
  ggplot(aes(x = test_alt, y = size_board_asl, fill = method)) +
    geom_boxplot() +
    geom_hline(yintercept = 2.93, linetype = 2, color = "red")
ggsave("plots/calibration_tests/LL_surface_board_asl.png", width = 10)
```

# Comparing Love Lab to BOI drones
```{r}
LL_vs_BOI_dat <- bind_rows(BOI_1_surface_dat, BOI_2_surface_dat, LL_surface_dat, .id = "drone")
View(LL_vs_BOI_dat)

LL_vs_BOI_dat %>% 
  filter(method == "video") %>%
  ggplot(aes(x = test_alt, y = size_pvc_asl, fill = drone)) +
    geom_boxplot() +
    geom_hline(yintercept = 1.98, linetype = 2, color = "red")
ggsave("plots/calibration_tests/LL_vs_BOI_pvc_asl.png", width = 10)

LL_vs_BOI_dat %>% 
  filter(method == "video") %>%
  ggplot(aes(x = test_alt, y = size_board_asl, fill = drone)) +
    geom_boxplot() +
    geom_hline(yintercept = 2.93, linetype = 2, color = "red")
ggsave("plots/calibration_tests/LL_vs_BOI_board_asl.png", width = 10)
```


Determining adjustment factor for Love Lab drone
```{r}
BOI_1_surface_dat %>% 
  filter(method == "video") %>% 
  group_by(test_alt) %>% 
  summarize(mean_pvc_asl = mean(size_pvc_asl), mean_board_asl = mean(size_board_asl)) %>% 
  mutate(error_pvc = (mean_pvc_asl/1.98)-1, error_board = (mean_board_asl/2.93)-1) %>% 
  gt()

BOI_2_surface_dat %>% 
  filter(method == "video") %>% 
  group_by(test_alt) %>% 
  summarize(mean_pvc_asl = mean(size_pvc_asl), mean_board_asl = mean(size_board_asl)) %>% 
  mutate(error_pvc = (mean_pvc_asl/1.98)-1, error_board = (mean_board_asl/2.93)-1) %>% 
  gt()

LL_surface_dat %>% 
  filter(method == "video") %>% 
  group_by(test_alt) %>% 
  summarize(mean_pvc_asl = mean(size_pvc_asl), mean_board_asl = mean(size_board_asl)) %>% 
  mutate(error_pvc = (mean_pvc_asl/1.98)-1, error_board = (mean_board_asl/2.93)-1) %>% 
  gt()

LL_surface_dat %>% 
  filter(method == "video") %>% 
  group_by(test_alt) %>% 
  summarize(mean_pvc_asl = (mean(size_pvc_asl))*0.72, mean_board_asl = mean(size_board_asl)*0.72) %>% 
  mutate(error_pvc = (mean_pvc_asl/1.98)-1, error_board = (mean_board_asl/2.93)-1) %>% 
  gt()

LL_surface_dat %>% 
  filter(method == "video") %>% 
  group_by(test_alt) %>% 
  summarize(mean_pvc_asl = mean(size_pvc_asl), mean_board_asl = mean(size_board_asl),
            mean_pvc_adj = mean(size_pvc_asl)*0.72, mean_board_adj = mean(size_board_asl)*0.72) %>% 
  pivot_longer(2:5, names_to = "type", values_to = "size") %>% 
  ggplot(aes(x = test_alt, y = size, color = type)) +
    geom_point(size = 4) +
    geom_hline(yintercept = 1.98, linetype = 2, color = "red") +
    geom_hline(yintercept = 2.93, linetype = 2, color = "red")
```

# AUV tests
```{r}
AUV.dat <- read.csv("data_raw/JP_measurements.csv")
View(AUV.dat)

AUV.dat <- mutate(AUV.dat, raw = if_else(is.na(yellow), full, yellow))
```

## Uncorrected estimates
```{r}
AUV.dat %>% 
  ggplot(aes(x = SRT_alt, y = raw)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = c(1.725, 2.4)) +
    facet_grid(cols =  vars(depth))
ggsave("plots/SRT_vs_raw_estimate.png")
```
## Error, corrected for depth of target
```{r}
AUV.dat %>% 
  ggplot(aes(x = SRT_alt, y = error)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 1) +
    facet_grid(cols = vars(depth))
ggsave("plots/SRT_vs_errors.png")
```

## Error, corrected for depth of target and using nominal test altitude (AGL + 0.77 m)
```{r}
AUV.dat %>% 
  ggplot(aes(x = test_alt, y = AGL_error)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 1) +
    facet_grid(cols = vars(depth)) +
    theme_bw()
ggsave("plots/AGL_vs_error.png")
```